---
title: "QC Report FluWarnSystem `r Sys.Date()`"
author: "FluWarnSystem"
date: "`r Sys.Date()`"
output: html_document
params:
  fasta: ""
  qc_table: "qc_table.tsv"
  subtype: ""
---

<style>
    body .main-container {
        max-width: 1800px;
    }
</style>



#### Table of contents

- [Overview](#item-zero)

- [Processed Sequences](#item-one)

- [Skipped Sequences](#item-two)


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```

```{r import, include=FALSE}
library(dplyr)

library(plotly)

library(DT)
```

```{r data, include=FALSE}
qc <- read.delim(params$qc_table)
qc <- qc %>% select(-X)

seq <- qc[is.na(qc$errors),]
row.names(seq) <- NULL

skipped <- qc[!is.na(qc$errors),]
row.names(skipped) <- NULL
```


<a id="item-zero"></a>

## Overview

The report was generated for the following input fasta file:

```{r input, echo=FALSE, message=FALSE, warning=FALSE}
cat(paste0('> ', params$fasta))
```

The sequences are from the subtype:

```{r subtype, echo=FALSE, message=FALSE, warning=FALSE}
if (params$subtype == "h1n1"){
  cat('> Influenza A H1N1pdm')
} else if (params$subtype == "h3n2"){
  cat('> Influenza A H3N2')
} else if (params$subtype == "vic"){
  cat('> Influenza B Victoria')
}
```

**Sequences that are not aligned were skipped because they were thrown out by Nextclade due to errors.**
For more information, please check the table in the section [Skipped Sequences](#item-two).

```{r pie chart, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
labels <- c('Sequences processed', 'Sequences skipped')
values <- c(nrow(seq), nrow(skipped))

plot_ly(labels = labels, values = values, type = "pie", marker = list(colors = c("darkgreen", "darkred"))) %>%
  layout(showlegend = TRUE)
```


<a id="item-one"></a>

## Processed Sequences {.tabset .tabset-fade}

### Table

```{r seq table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(seq) == 0){
  cat("All sequences were skipped")
} else {
  seq_tab <- seq %>%
    select(ID, qc.overallStatus, alignmentScore, coverage, failedCdses, warnings)
  
  names(seq_tab)[2] <- "QC Status"
  
  names(seq_tab)[3] <- "Alignment Score"
  names(seq_tab)[4] <- "Coverage"
  seq_tab$Coverage <- round(seq_tab$Coverage, 2)
  
  names(seq_tab)[5] <- "Failed"
  names(seq_tab)[6] <- "Warnings"
  
  datatable(
    seq_tab, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 10,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )
}
```

### QC Status

```{r qc status, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(seq) == 0){
  cat("All sequences were skipped")
} else {  
  bar.data <- seq %>%
    group_by(qc.overallStatus) %>%
    summarize(count = n())
  
  status_tick <- unique(seq$qc.overallStatus)
  
  plot_ly(
      bar.data, x=~qc.overallStatus, y=~count, 
      type = "bar",
      marker = list(color = "lightblue"),
      hoverinfo='text',
      hovertext=~paste0("<b>Status: </b>", .data$qc.overallStatus, "<br><b># of Sequences: </b>", .data$count)
    ) %>%
      layout(
        xaxis=list(nticks=length(status_tick), title='<b>QC Status</b>', showgrid=FALSE),
        yaxis=list(title="<b>Number of Sequences</b>", showgrid=FALSE)
      )
}
```

### Alignment score

```{r alignment, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(seq) == 0){
  cat("All sequences were skipped")
} else {
  align.data <- seq %>%
    group_by(alignmentScore) %>%
    summarize(count = n())
  
  plot_ly(
    align.data, x = ~alignmentScore, y = ~count,
    type = "bar",
    marker = list(color = "blue"),
    hoverinfo='text',
    hovertext=~paste0("<b>Alignment Score: </b>", .data$alignmentScore, "<br><b># of Sequences: </b>", .data$count)
  ) %>%
    layout(
      xaxis = list(title = "Alignment Score"),
      yaxis = list(title = "Number of Sequences")
    )
}
```

### Coverage

```{r coverage, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(seq) == 0){
  cat("All sequences were skipped")
} else {  
  cov.data <- seq_tab %>%
    group_by(Coverage) %>%
    summarize(count = n())
  
  plot_ly(
    cov.data, x = ~Coverage, y = ~count,
    type = "bar",
    marker = list(color = "blue"),
    hoverinfo='text',
    hovertext=~paste0("<b>Coverage: </b>", .data$Coverage, "<br><b># of Sequences: </b>", .data$count)
  ) %>%
    layout(
      yaxis = list(title = "Number of Sequences")
    )
}
```


<a id="item-two"></a>

## Skipped Sequences 

```{r skipped seq table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(skipped) == 0){
  cat("No sequences were skipped")
} else {
  skipped_tab <- skipped %>%
    select(ID, errors)
  
  names(skipped_tab)[2] <- "Error"
  
  datatable(
    skipped_tab, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 10,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )
}
```