---
title: "Report FluWarnSystem `r Sys.Date()`"
author: "FluWarnSystem"
date: "`r Sys.Date()`"
output: html_document
params:
  fasta: ""
  metadata: ""
  variant_table: "variant_table.tsv"
  alert_samples: "fluwarnsystem-alerts-samples-all.csv"
  subtype: ""
  ref: ""
  moc: ""
  roi: ""
---

<style>
    body .main-container {
        max-width: 1800px;
    }
</style>




#### Table of contents

- [Overview](#item-zero)

- [Group RED](#item-one)

- [Group ORANGE](#item-two)

- [Group YELLOW](#item-three)

- [Group GREY](#item-four)

- [Group BLACK](#item-five)

- [Reference sequences](#item-six)

- [MOC table](#item-seven)

- [ROI table](#item-eight)




#### List of Abbrevations

Sequences are sent to a category (Red, Orange, Yellow, Grey, Black) according to their mutation profile. FluWarnSystem counts the amount of mutations in 3 categories:

**MOC - Mutation of Concern**

Those are mutations that have been identified previously in the literature as changing the potential of the virus. The table of all MOC with their bibliographical reference is at the bottom in the [MOC table](#item-seven).

**ROI - Region of Interest**

Those are important regions of the gene like for example Antigenic sites or Receptor-binding sites. The table of all ROI with their bibliographical reference is at the bottom in the [ROI table](#item-eight).

**PM - Private Mutation**

Mutations that are neither a MOC nor a ROI.


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```

```{r import, include=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
library(stringi)
library(readxl)

library(seqinr)

library(ggplot2)
library(plotly)

library(DT)
```

```{r data, include=FALSE}
fasta <- read.fasta(params$fasta)

variant_table <- read.delim(params$variant_table)
samples <- read.delim(params$alert_samples, sep = ',')

moc <- read.delim(params$moc)
roi <- read.delim(params$roi, sep = ',')

if (params$metadata != ""){
  metadata <- read_xlsx(params$metadata)
}
```

```{r remove, include=FALSE}
samples <- samples %>%
  filter(!ID %in% c('ROI_ERROR_CATCH', 'NotAnnotated_ERROR_CATCH'))
```


<a id="item-zero"></a>

## Overview

The report was generated for the following input fasta file:

```{r input, echo=FALSE, message=FALSE, warning=FALSE}
cat(paste0('> ', params$fasta))
```

The sequences are from the subtype:

```{r subtype, echo=FALSE, message=FALSE, warning=FALSE}
if (params$subtype == "h1n1"){
  cat('> Influenza A H1N1pdm')
} else if (params$subtype == "h3n2"){
  cat('> Influenza A H3N2')
} else if (params$subtype == "vic"){
  cat('> Influenza B Victoria')
}
```

**Sequences that are not aligned were skipped because they were thrown out by Nextclade due to errors.**
For more information, please check the table in the section Skipped Sequences in the QC Report.

```{r seq total, echo=FALSE, message=FALSE, warning=FALSE}
cat(paste0('> Sequences total in fasta file: ', length(fasta)))
cat(paste0('> Sequences aligned: ', nrow(samples)))
cat(paste0('> Sequences skipped: ', length(fasta) - nrow(samples)))
```

```{r metadata, echo=FALSE, message=FALSE, warning=FALSE}
if (params$metadata != ""){
  cat(paste0('> Metadata was given for the following amount of sequences: ', nrow(metadata)))
}
```


<a id="item-one"></a>

## Group <span style="color: red;">RED</span> {.tabset .tabset-fade}
#### *Sequences with at least 1 MOC and 5 or more MOC & ROI*

```{r red data, echo=FALSE, message=FALSE, warning=FALSE}
red_table <- filter(samples, alert_level == 'red') %>%
  select(ID, ListMutationsSelected_M, ListMutationsSelected_D, ListMutationsSelected_I)

for (i in red_table["ID"]) {
  new <- str_remove_all(i, "%space%")
  red_table["ID"][red_table["ID"] == i] <- new
}

names(red_table)[2] <- "Mutations"
names(red_table)[3] <- "Deletions"
names(red_table)[4] <- "Insertions"

red_table_copy <- red_table

for (i in red_table["Mutations"]){
  new <- str_split(i, ",")
  red_table["Mutations"][red_table["Mutations"] == i] <- new
}

cat(paste0('> ', 'Sequences in Group RED: ', nrow(red_table)))
```

### Lollipop Plot

```{r red lollipop, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(red_table) == 0){
  cat("No sequences in Group RED")
} else {
  clade <- c()
  for(i in red_table$ID){
    new_clade <- variant_table[variant_table$ID == i, "clade"]
    clade <- c(clade, new_clade[[1]])
  }
  
  clade_df <- red_table
  clade_df$clade <- clade
  clade_df$group <- "All Sequences"
  clade_unique <- unique(clade)
  clade_len <- length(unique(clade))
  
  df_list <- list(clade_df)
  for (i in clade_unique){
    new_df <- subset(clade_df, clade == i)
    new_df$group <- i
    df_list <- c(df_list, list(new_df))
  }
  
  groups <- c()
  data_list <- list()
  for (df in df_list){
    mut_freq <- rep(0, 566)
    mut_list <- vector(mode = "character", length = 566)
    mut_list_unique <- vector(mode = "character", length = 566)
    mut_list_unique_len <- vector(mode = "integer", length = 566)
    for (i in df["Mutations"]){
      for(j in i){
        # get number of mutations at every position
        ind <- as.integer(str_sub(j, 2, -2))
        mut_freq[ind] <- mut_freq[ind] + 1
        # get list of mutations at every position
        mut_list[ind] <- paste(j, mut_list[ind], sep=" ")
      }
    }
    
    pos <- 1
    for (i in mut_list){
      split <- str_split_1(i, " ")
      p <- 1
      for (k in split){
        if (grepl("\\*", k) == TRUE){
          split[p] <- paste("\\", k, sep = "")
          print(split[k])
        }
        p <- p + 1
      }
      unique_mut <- stri_unique(split)
      for (j in unique_mut[-length(unique_mut)]){
        count <- str_count(i, j)
        mut_list_unique[pos] <- paste(mut_list_unique[pos], "<br>", toString(count), j)
        mut_list_unique_len[pos] <- mut_list_unique_len[pos] + 1
      }
      pos <- pos + 1
    }
    
    data_temp <- data.frame(x = seq(1, 566), y = mut_freq, list = mut_list_unique, list_len = mut_list_unique_len, group = df$group[1])
    groups <- c(groups, df$group[1])
    data_list <- c(data_list, list(data_temp))
  }
  
  all_false <- rep(FALSE, length(groups)*2)
  button_list <- list()
  for (i in 1:length(groups)) {
    visibility_state <- all_false
    visibility_state[(2 * i - 1):(i + i)] <- TRUE
    button_list[[i]] <- list(
      label = groups[i],
      method = "update",
      args = list(list(visible = visibility_state), list(title = groups[i]))
    )
  }
  
  p <- ggplot() +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank()
    ) +
    xlab("Amino Acid Position") +
    ylab("Number of samples") +
    xlim(c(0,566))
  
  for (i in 1:length(groups)) {
    d <- data_list[[i]][which(data_list[[i]]$y > 0),]
    row.names(d) <- seq(nrow(d))
    p <- p + 
      geom_segment(data = d, aes(x=x, xend=x, y=0, yend=y), color="darkgrey") +
      geom_point(
        data = d, 
        aes(
          x=x, y=y, 
          text=paste0(
          "<b>AA Position: </b>", x, 
          "<br><b># of alternative AA: </b>", y, 
          "<br><b>Substitutions: </b>", list_len, list
          )
        ),
        color="red",
        size=3
      )
  }
  
  ggplotly(p, tooltip = c("text")) %>% layout(
    updatemenus=list(
      list(
        active=0, 
        x=0, 
        y=1.1,
        xanchor="left",
        yanchor="bottom",
        buttons=button_list
      )
    )
  )
}
```

### Table

```{r red table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(red_table) == 0){
  cat("No sequences in Group RED")
} else {
  for (i in red_table_copy["Mutations"]){
    new <- gsub(",([[:alpha:]])", ", \\1", i)
    red_table_copy["Mutations"][red_table_copy["Mutations"] == i] <- new
  }
  
  subtype <- vector(mode = "character", length = length(red_table_copy$ID))
  seg <- vector(mode = "character", length = length(red_table_copy$ID))
    
  red_table_copy <- cbind(subtype, seg, red_table_copy)
    
  names(red_table_copy)[1] <- "Subtype"
  names(red_table_copy)[2] <- "Segment"
  
  for (j in 1:length(red_table_copy$ID)){
    red_table_copy$Subtype[j] <- str_split_i(red_table_copy$ID[j], "\\|", -1)
    red_table_copy$Segment[j] <- str_split_i(red_table_copy$ID[j], "\\|", 2)
  }
  
  datatable(
    red_table_copy, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 5,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      columnDefs = list(list(width = 500, targets = 4)),
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )
}
```

### Heatmap

```{r red heatmap, out.width="1200px", out.height="800px", echo=FALSE, message=FALSE, warning=FALSE}
if (!(nrow(red_table) == 0)){
  names <- vector(mode = "character", length = length(red_table$ID))
  df_names <- cbind(red_table, names)
  names(df_names)[5] <- "Isolate_Name"
  for (j in 1:length(df_names$ID)){
    df_names$Isolate_Name[j] <- trimws(str_split_i(df_names$ID[j], "\\|", 3))
  }
}

if (nrow(red_table) == 0){
  cat("No sequences in Group RED")
} else if (params$metadata == ""){
  cat("The heatmap shows the frequency of mutations over the weeks.")
  cat("No metadata file with collection dates was given, so no heatmap was generated.")
} else if (any(df_names$Isolate_Name %in% metadata$Isolate_Name) == FALSE){
  cat("No metadata given for the sequences of Group RED.")
} else {
  df_mutations_filtered <- merge(metadata, df_names, by="Isolate_Name") %>%
    select(Collection_Date, Mutations) %>%
    unnest(Mutations) %>%
    mutate(week = week(ymd(Collection_Date)))
    
  df_week <- df_mutations_filtered %>% 
    group_by(week) %>%
    summarize(count = n())
  
  df_mutations_filtered <- merge(df_mutations_filtered, df_week, by="week") %>%
    select(Mutations, week, count)

  names(df_mutations_filtered)[1] <- "aa_profile"
  names(df_mutations_filtered)[3] <- "week_count_total"

  df_mutations_grouped <- df_mutations_filtered %>% 
    group_by(week, week_count_total, aa_profile) %>% 
    summarize(freq = n()) %>%
    mutate(perc = round(freq/week_count_total*100, digits = 2)) %>% 
    ungroup() 
  
  pos <- c()
  for (i in df_mutations_grouped["aa_profile"]){
    # get position of mutation
    ind <- as.integer(str_sub(i, 2, -2))
    pos <- c(pos, ind)
  }
  df_mutations_grouped$position <- pos
  
  df_mutations_sorted <- df_mutations_grouped %>%
    arrange(-position)
  
  plot.data <- df_mutations_sorted %>% 
    select(week, aa_profile, freq, perc)
  
  min_val <- min(plot.data$week)
  max_val <- max(plot.data$week)
  full_seq <- min_val:max_val
  missing <- setdiff(full_seq, plot.data$week)
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], aa_profile = NA, freq = NA, perc = NA)
    plot.data <- rbind(plot.data, new_row)
  }
  
  aa_tick <- unique(plot.data$aa_profile)
  week_tick <- unique(plot.data$week)
  
  heatmap <- plot_ly(
    plot.data, x=~week, y=~aa_profile, z=~freq, 
    reversescale=TRUE, 
    xgap=1.5, ygap=1.5, 
    type="heatmap", 
    colorbar=list(
      title=('<b>Freq</b>')), 
      hoverinfo='text', 
      text=~paste0("<b>Week: </b>", .data$week, "<br><b>Substitution: </b>", .data$aa_profile, "<br><b># of Substitutions: </b>", .data$freq)
    ) %>% 
    layout(
      plot_bgcolor="#F8F8F8", 
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(nticks=length(aa_tick), title="<b>Substitutions</b>", showgrid=FALSE, categoryorder="trace"),
      legend=list(orientation='v', xanchor="right", yanchor="top")
    ) 
  
  bar.data <- metadata %>%
    select(Collection_Date) %>%
    mutate(week = week(ymd(Collection_Date))) %>%
    group_by(week) %>%
    summarize(count = n())
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], count = 0)
    bar.data <- rbind(bar.data, new_row)
  }
  
  bar.data <- bar.data %>% 
    arrange(week)
  
  barplot <- plot_ly(
    bar.data, x=~week, y=~count, 
    type = "bar",
    marker = list(color = "grey"),
    hoverinfo='text',
    hovertext=~paste0("<b>Week: </b>", .data$week, "<br><b># of Sequences: </b>", .data$count)
  ) %>%
    layout(
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(title="<b>Total Number <br> of Sequences</b>", showgrid=FALSE)
    )
  
  subplot(
    heatmap, barplot, 
    nrows = 2, heights = c(0.8, 0.2), 
    shareX = TRUE, titleY = TRUE
  )
}
```


<a id="item-two"></a>

## Group <span style="color: orange;">ORANGE</span> {.tabset .tabset-fade}
#### *Sequences with at least 1 MOC and 3 or more MOC & ROI*

```{r orange data, echo=FALSE, message=FALSE, warning=FALSE}
orange_table <- filter(samples, alert_level == 'orange') %>%
  select(ID, ListMutationsSelected_M, ListMutationsSelected_D, ListMutationsSelected_I)

for (i in orange_table["ID"]) {
  new <- str_remove_all(i, "%space%")
  orange_table["ID"][orange_table["ID"] == i] <- new
}

names(orange_table)[2] <- "Mutations"
names(orange_table)[3] <- "Deletions"
names(orange_table)[4] <- "Insertions"

orange_table_copy <- orange_table

for (i in orange_table["Mutations"]){
  new <- str_split(i, ",")
  orange_table["Mutations"][orange_table["Mutations"] == i] <- new
}

cat(paste0('> ', 'Sequences in Group ORANGE: ', nrow(orange_table)))
```

### Lollipop Plot

```{r orange lollipop, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(orange_table) == 0){
  cat("No sequences in Group ORANGE")
} else {
  clade <- c()
  for(i in orange_table$ID){
    new_clade <- variant_table[variant_table$ID == i, "clade"]
    clade <- c(clade, new_clade[[1]])
  }
  
  clade_df <- orange_table
  clade_df$clade <- clade
  clade_df$group <- "All Sequences"
  clade_unique <- unique(clade)
  clade_len <- length(unique(clade))
  
  df_list <- list(clade_df)
  for (i in clade_unique){
    new_df <- subset(clade_df, clade == i)
    new_df$group <- i
    df_list <- c(df_list, list(new_df))
  }
  
  groups <- c()
  data_list <- list()
  for (df in df_list){
    mut_freq <- rep(0, 566)
    mut_list <- vector(mode = "character", length = 566)
    mut_list_unique <- vector(mode = "character", length = 566)
    mut_list_unique_len <- vector(mode = "integer", length = 566)
    for (i in df["Mutations"]){
      for(j in i){
        # get number of mutations at every position
        ind <- as.integer(str_sub(j, 2, -2))
        mut_freq[ind] <- mut_freq[ind] + 1
        # get list of mutations at every position
        mut_list[ind] <- paste(j, mut_list[ind], sep=" ")
      }
    }
    
    pos <- 1
    for (i in mut_list){
      split <- str_split_1(i, " ")
      p <- 1
      for (k in split){
        if (grepl("\\*", k) == TRUE){
          split[p] <- paste("\\", k, sep = "")
          print(split[k])
        }
        p <- p + 1
      }
      unique_mut <- stri_unique(split)
      for (j in unique_mut[-length(unique_mut)]){
        count <- str_count(i, j)
        mut_list_unique[pos] <- paste(mut_list_unique[pos], "<br>", toString(count), j)
        mut_list_unique_len[pos] <- mut_list_unique_len[pos] + 1
      }
      pos <- pos + 1
    }
    
    data_temp <- data.frame(x = seq(1, 566), y = mut_freq, list = mut_list_unique, list_len = mut_list_unique_len, group = df$group[1])
    groups <- c(groups, df$group[1])
    data_list <- c(data_list, list(data_temp))
  }
  
  all_false <- rep(FALSE, length(groups)*2)
  button_list <- list()
  for (i in 1:length(groups)) {
    visibility_state <- all_false
    visibility_state[(2 * i - 1):(i + i)] <- TRUE
    button_list[[i]] <- list(
      label = groups[i],
      method = "update",
      args = list(list(visible = visibility_state), list(title = groups[i]))
    )
  }
  
  p <- ggplot() +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank()
    ) +
    xlab("Amino Acid Position") +
    ylab("Number of samples") +
    xlim(c(0,566))
  
  for (i in 1:length(groups)) {
    d <- data_list[[i]][which(data_list[[i]]$y > 0),]
    row.names(d) <- seq(nrow(d))
    p <- p + 
      geom_segment(data = d, aes(x=x, xend=x, y=0, yend=y), color="darkgrey") +
      geom_point(
        data = d, 
        aes(
          x=x, y=y, 
          text=paste0(
          "<b>AA Position: </b>", x, 
          "<br><b># of alternative AA: </b>", y, 
          "<br><b>Substitutions: </b>", list_len, list
          )
        ),
        color="orange",
        size=3
      )
  }
  
  ggplotly(p, tooltip = c("text")) %>% layout(
    updatemenus=list(
      list(
        active=0, 
        x=0, 
        y=1.1,
        xanchor="left",
        yanchor="bottom",
        buttons=button_list
      )
    )
  )
}
```

### Table

```{r orange table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(orange_table) == 0){
  cat("No sequences in Group ORANGE")
} else {
  for (i in orange_table_copy["Mutations"]){
    new <- gsub(",([[:alpha:]])", ", \\1", i)
    orange_table_copy["Mutations"][orange_table_copy["Mutations"] == i] <- new
  }
  
  subtype <- vector(mode = "character", length = length(orange_table_copy$ID))
  seg <- vector(mode = "character", length = length(orange_table_copy$ID))
    
  orange_table_copy <- cbind(subtype, seg, orange_table_copy)
    
  names(orange_table_copy)[1] <- "Subtype"
  names(orange_table_copy)[2] <- "Segment"
  
  for (j in 1:length(orange_table_copy$ID)){
    orange_table_copy$Subtype[j] <- str_split_i(orange_table_copy$ID[j], "\\|", -1)
    orange_table_copy$Segment[j] <- str_split_i(orange_table_copy$ID[j], "\\|", 2)
  }
  
  datatable(
    orange_table_copy, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 5,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      columnDefs = list(list(width = 500, targets = 4)),
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )
}
```

### Heatmap

```{r orange heatmap, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (!(nrow(orange_table) == 0)){
  names <- vector(mode = "character", length = length(orange_table$ID))
  df_names <- cbind(orange_table, names)
  names(df_names)[5] <- "Isolate_Name"
  for (j in 1:length(df_names$ID)){
    df_names$Isolate_Name[j] <- trimws(str_split_i(df_names$ID[j], "\\|", 3))
  }
}

if (nrow(orange_table) == 0){
  cat("No sequences in Group ORANGE")
} else if (params$metadata == ""){
  cat("The heatmap shows the frequency of mutations over the weeks.")
  cat("No metadata file with collection dates was given, so no heatmap was generated.")
} else if (any(df_names$Isolate_Name %in% metadata$Isolate_Name) == FALSE){
  cat("No metadata given for the sequences of Group ORANGE.")
} else {
  df_mutations_filtered <- merge(metadata, df_names, by="Isolate_Name") %>%
    select(Collection_Date, Mutations) %>%
    unnest(Mutations) %>%
    mutate(week = week(ymd(Collection_Date)))
    
  df_week <- df_mutations_filtered %>% 
    group_by(week) %>%
    summarize(count = n())
  
  df_mutations_filtered <- merge(df_mutations_filtered, df_week, by="week") %>%
    select(Mutations, week, count)

  names(df_mutations_filtered)[1] <- "aa_profile"
  names(df_mutations_filtered)[3] <- "week_count_total"

  df_mutations_grouped <- df_mutations_filtered %>% 
    group_by(week, week_count_total, aa_profile) %>% 
    summarize(freq = n()) %>%
    mutate(perc = round(freq/week_count_total*100, digits = 2)) %>% 
    ungroup() 
  
  pos <- c()
  for (i in df_mutations_grouped["aa_profile"]){
    # get position of mutation
    ind <- as.integer(str_sub(i, 2, -2))
    pos <- c(pos, ind)
  }
  df_mutations_grouped$position <- pos
  
  df_mutations_sorted <- df_mutations_grouped %>%
    arrange(-position)
  
  plot.data <- df_mutations_sorted %>% 
    select(week, aa_profile, freq, perc)
  
  min_val <- min(plot.data$week)
  max_val <- max(plot.data$week)
  full_seq <- min_val:max_val
  missing <- setdiff(full_seq, plot.data$week)
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], aa_profile = NA, freq = NA, perc = NA)
    plot.data <- rbind(plot.data, new_row)
  }
  
  aa_tick <- unique(plot.data$aa_profile)
  week_tick <- unique(plot.data$week)
  
  heatmap <- plot_ly(
    plot.data, x=~week, y=~aa_profile, z=~freq, 
    reversescale=TRUE, 
    xgap=1.5, ygap=1.5, 
    type="heatmap", 
    colorbar=list(
      title=('<b>Freq</b>')), 
      hoverinfo='text', 
      text=~paste0("<b>Week: </b>", .data$week, "<br><b>Substitution: </b>", .data$aa_profile, "<br><b># of Substitutions: </b>", .data$freq)
    ) %>% 
    layout(
      plot_bgcolor="#F8F8F8", 
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(nticks=length(aa_tick), title="<b>Substitutions</b>", showgrid=FALSE, categoryorder="trace"),
      legend=list(orientation='v', xanchor="right", yanchor="top")
    ) 
  
  bar.data <- metadata %>%
    select(Collection_Date) %>%
    mutate(week = week(ymd(Collection_Date))) %>%
    group_by(week) %>%
    summarize(count = n())
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], count = 0)
    bar.data <- rbind(bar.data, new_row)
  }
  
  bar.data <- bar.data %>% 
    arrange(week)
  
  barplot <- plot_ly(
    bar.data, x=~week, y=~count, 
    type = "bar",
    marker = list(color = "grey"),
    hoverinfo='text',
    hovertext=~paste0("<b>Week: </b>", .data$week, "<br><b># of Sequences: </b>", .data$count)
  ) %>%
    layout(
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(title="<b>Total Number <br> of Sequences</b>", showgrid=FALSE)
    )
  
  subplot(
    heatmap, barplot, 
    nrows = 2, heights = c(0.8, 0.2), 
    shareX = TRUE, titleY = TRUE
  )
}
```


<a id="item-three"></a>

## Group <span style="color: #FFEA00;">YELLOW</span> {.tabset .tabset-fade}
#### *Sequences with 0 MOC and 25 or more PM and 7 or less ROI*
#### *Sequences with 0 MOC and 8 or more ROI and 24 or less PM*

```{r yellow data, echo=FALSE, message=FALSE, warning=FALSE}
yellow_table <- filter(samples, alert_level == '#FFEA00') %>%
  select(ID, ListMutationsSelected_M, ListMutationsSelected_D, ListMutationsSelected_I)

for (i in yellow_table["ID"]) {
  new <- str_remove_all(i, "%space%")
  yellow_table["ID"][yellow_table["ID"] == i] <- new
}

names(yellow_table)[2] <- "Mutations"
names(yellow_table)[3] <- "Deletions"
names(yellow_table)[4] <- "Insertions"

yellow_table_copy <- yellow_table

for (i in yellow_table["Mutations"]){
  new <- str_split(i, ",")
  yellow_table["Mutations"][yellow_table["Mutations"] == i] <- new
}

cat(paste0('> ', 'Sequences in Group YELLOW: ', nrow(yellow_table)))
```

### Lollipop Plot

```{r yellow lollipop, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(yellow_table) == 0){
  cat("No sequences in Group YELLOW")
} else {
  clade <- c()
  for(i in yellow_table$ID){
    new_clade <- variant_table[variant_table$ID == i, "clade"]
    clade <- c(clade, new_clade[[1]])
  }
  
  clade_df <- yellow_table
  clade_df$clade <- clade
  clade_df$group <- "All Sequences"
  clade_unique <- unique(clade)
  clade_len <- length(unique(clade))
  
  df_list <- list(clade_df)
  for (i in clade_unique){
    new_df <- subset(clade_df, clade == i)
    new_df$group <- i
    df_list <- c(df_list, list(new_df))
  }
  
  groups <- c()
  data_list <- list()
  for (df in df_list){
    mut_freq <- rep(0, 566)
    mut_list <- vector(mode = "character", length = 566)
    mut_list_unique <- vector(mode = "character", length = 566)
    mut_list_unique_len <- vector(mode = "integer", length = 566)
    for (i in df["Mutations"]){
      for(j in i){
        # get number of mutations at every position
        ind <- as.integer(str_sub(j, 2, -2))
        mut_freq[ind] <- mut_freq[ind] + 1
        # get list of mutations at every position
        mut_list[ind] <- paste(j, mut_list[ind], sep=" ")
      }
    }
    
    pos <- 1
    for (i in mut_list){
      split <- str_split_1(i, " ")
      p <- 1
      for (k in split){
        if (grepl("\\*", k) == TRUE){
          split[p] <- paste("\\", k, sep = "")
          print(split[k])
        }
        p <- p + 1
      }
      unique_mut <- stri_unique(split)
      for (j in unique_mut[-length(unique_mut)]){
        count <- str_count(i, j)
        mut_list_unique[pos] <- paste(mut_list_unique[pos], "<br>", toString(count), j)
        mut_list_unique_len[pos] <- mut_list_unique_len[pos] + 1
      }
      pos <- pos + 1
    }
    
    data_temp <- data.frame(x = seq(1, 566), y = mut_freq, list = mut_list_unique, list_len = mut_list_unique_len, group = df$group[1])
    groups <- c(groups, df$group[1])
    data_list <- c(data_list, list(data_temp))
  }
  
  all_false <- rep(FALSE, length(groups)*2)
  button_list <- list()
  for (i in 1:length(groups)) {
    visibility_state <- all_false
    visibility_state[(2 * i - 1):(i + i)] <- TRUE
    button_list[[i]] <- list(
      label = groups[i],
      method = "update",
      args = list(list(visible = visibility_state), list(title = groups[i]))
    )
  }
  
  p <- ggplot() +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank()
    ) +
    xlab("Amino Acid Position") +
    ylab("Number of samples") +
    xlim(c(0,566))
  
  for (i in 1:length(groups)) {
    d <- data_list[[i]][which(data_list[[i]]$y > 0),]
    row.names(d) <- seq(nrow(d))
    p <- p + 
      geom_segment(data = d, aes(x=x, xend=x, y=0, yend=y), color="darkgrey") +
      geom_point(
        data = d, 
        aes(
          x=x, y=y, 
          text=paste0(
          "<b>AA Position: </b>", x, 
          "<br><b># of alternative AA: </b>", y, 
          "<br><b>Substitutions: </b>", list_len, list
          )
        ),
        color="#FFEA00",
        size=3
      )
  }
  
  ggplotly(p, tooltip = c("text")) %>% layout(
    updatemenus=list(
      list(
        active=0, 
        x=0, 
        y=1.1,
        xanchor="left",
        yanchor="bottom",
        buttons=button_list
      )
    )
  )
}
```

### Table

```{r yellow table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(yellow_table) == 0){
  cat("No sequences in Group YELLOW")
} else {
  for (i in yellow_table_copy["Mutations"]){
    new <- gsub(",([[:alpha:]])", ", \\1", i)
    yellow_table_copy["Mutations"][yellow_table_copy["Mutations"] == i] <- new
  }
  
  subtype <- vector(mode = "character", length = length(yellow_table_copy$ID))
  seg <- vector(mode = "character", length = length(yellow_table_copy$ID))
    
  yellow_table_copy <- cbind(subtype, seg, yellow_table_copy)
    
  names(yellow_table_copy)[1] <- "Subtype"
  names(yellow_table_copy)[2] <- "Segment"
  
  for (j in 1:length(yellow_table_copy$ID)){
    yellow_table_copy$Subtype[j] <- str_split_i(yellow_table_copy$ID[j], "\\|", -1)
    yellow_table_copy$Segment[j] <- str_split_i(yellow_table_copy$ID[j], "\\|", 2)
  }
  
  datatable(
    yellow_table_copy, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 5,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      columnDefs = list(list(width = 500, targets = 4)),
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )
}
```

### Heatmap

```{r yellow heatmap, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (!(nrow(yellow_table) == 0)){
  names <- vector(mode = "character", length = length(yellow_table$ID))
  df_names <- cbind(yellow_table, names)
  names(df_names)[5] <- "Isolate_Name"
  for (j in 1:length(df_names$ID)){
    df_names$Isolate_Name[j] <- trimws(str_split_i(df_names$ID[j], "\\|", 3))
  }
}

if (nrow(yellow_table) == 0){
  cat("No sequences in Group YELLOW")
} else if (params$metadata == ""){
  cat("The heatmap shows the frequency of mutations over the weeks.")
  cat("No metadata file with collection dates was given, so no heatmap was generated.")
} else if (any(df_names$Isolate_Name %in% metadata$Isolate_Name) == FALSE){
  cat("No metadata given for the sequences of Group YELLOW.")
} else {
  df_mutations_filtered <- merge(metadata, df_names, by="Isolate_Name") %>%
    select(Collection_Date, Mutations) %>%
    unnest(Mutations) %>%
    mutate(week = week(ymd(Collection_Date)))
    
  df_week <- df_mutations_filtered %>% 
    group_by(week) %>%
    summarize(count = n())
  
  df_mutations_filtered <- merge(df_mutations_filtered, df_week, by="week") %>%
    select(Mutations, week, count)

  names(df_mutations_filtered)[1] <- "aa_profile"
  names(df_mutations_filtered)[3] <- "week_count_total"

  df_mutations_grouped <- df_mutations_filtered %>% 
    group_by(week, week_count_total, aa_profile) %>% 
    summarize(freq = n()) %>%
    mutate(perc = round(freq/week_count_total*100, digits = 2)) %>% 
    ungroup() 
  
  pos <- c()
  for (i in df_mutations_grouped["aa_profile"]){
    # get position of mutation
    ind <- as.integer(str_sub(i, 2, -2))
    pos <- c(pos, ind)
  }
  df_mutations_grouped$position <- pos
  
  df_mutations_sorted <- df_mutations_grouped %>%
    arrange(-position)
  
  plot.data <- df_mutations_sorted %>% 
    select(week, aa_profile, freq, perc)
  
  min_val <- min(plot.data$week)
  max_val <- max(plot.data$week)
  full_seq <- min_val:max_val
  missing <- setdiff(full_seq, plot.data$week)
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], aa_profile = NA, freq = NA, perc = NA)
    plot.data <- rbind(plot.data, new_row)
  }
  
  aa_tick <- unique(plot.data$aa_profile)
  week_tick <- unique(plot.data$week)
  
  heatmap <- plot_ly(
    plot.data, x=~week, y=~aa_profile, z=~freq, 
    reversescale=TRUE, 
    xgap=1.5, ygap=1.5, 
    type="heatmap", 
    colorbar=list(
      title=('<b>Freq</b>')), 
      hoverinfo='text', 
      text=~paste0("<b>Week: </b>", .data$week, "<br><b>Substitution: </b>", .data$aa_profile, "<br><b># of Substitutions: </b>", .data$freq)
    ) %>% 
    layout(
      plot_bgcolor="#F8F8F8", 
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(nticks=length(aa_tick), title="<b>Substitutions</b>", showgrid=FALSE, categoryorder="trace"),
      legend=list(orientation='v', xanchor="right", yanchor="top")
    ) 
  
  bar.data <- metadata %>%
    select(Collection_Date) %>%
    mutate(week = week(ymd(Collection_Date))) %>%
    group_by(week) %>%
    summarize(count = n())
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], count = 0)
    bar.data <- rbind(bar.data, new_row)
  }
  
  bar.data <- bar.data %>% 
    arrange(week)
  
  barplot <- plot_ly(
    bar.data, x=~week, y=~count, 
    type = "bar",
    marker = list(color = "grey"),
    hoverinfo='text',
    hovertext=~paste0("<b>Week: </b>", .data$week, "<br><b># of Sequences: </b>", .data$count)
  ) %>%
    layout(
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(title="<b>Total Number <br> of Sequences</b>", showgrid=FALSE)
    )
  
  subplot(
    heatmap, barplot, 
    nrows = 2, heights = c(0.8, 0.2), 
    shareX = TRUE, titleY = TRUE
  )
}
```


<a id="item-four"></a>

## Group <span style="color: lightgrey;">GREY</span> {.tabset .tabset-fade}
#### *Sequences with 25 or more PM and 8 or more ROI*

```{r grey data, echo=FALSE, message=FALSE, warning=FALSE}
grey_table <- filter(samples, alert_level == 'grey') %>%
  select(ID, ListMutationsSelected_M, ListMutationsSelected_D, ListMutationsSelected_I)

for (i in grey_table["ID"]) {
  new <- str_remove_all(i, "%space%")
  grey_table["ID"][grey_table["ID"] == i] <- new
}

names(grey_table)[2] <- "Mutations"
names(grey_table)[3] <- "Deletions"
names(grey_table)[4] <- "Insertions"

grey_table_copy <- grey_table

for (i in grey_table["Mutations"]){
  new <- str_split(i, ",")
  grey_table["Mutations"][grey_table["Mutations"] == i] <- new
}

cat(paste0('> ', 'Sequences in Group GREY: ', nrow(grey_table)))
```

### Lollipop Plot

```{r grey lollipop, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(grey_table) == 0){
  cat("No sequences in Group GREY")
} else {
  clade <- c()
  for(i in grey_table$ID){
    new_clade <- variant_table[variant_table$ID == i, "clade"]
    clade <- c(clade, new_clade[[1]])
  }
  
  clade_df <- grey_table
  clade_df$clade <- clade
  clade_df$group <- "All Sequences"
  clade_unique <- unique(clade)
  clade_len <- length(unique(clade))
  
  df_list <- list(clade_df)
  for (i in clade_unique){
    new_df <- subset(clade_df, clade == i)
    new_df$group <- i
    df_list <- c(df_list, list(new_df))
  }
  
  groups <- c()
  data_list <- list()
  for (df in df_list){
    mut_freq <- rep(0, 566)
    mut_list <- vector(mode = "character", length = 566)
    mut_list_unique <- vector(mode = "character", length = 566)
    mut_list_unique_len <- vector(mode = "integer", length = 566)
    for (i in df["Mutations"]){
      for(j in i){
        # get number of mutations at every position
        ind <- as.integer(str_sub(j, 2, -2))
        mut_freq[ind] <- mut_freq[ind] + 1
        # get list of mutations at every position
        mut_list[ind] <- paste(j, mut_list[ind], sep=" ")
      }
    }
    
    pos <- 1
    for (i in mut_list){
      split <- str_split_1(i, " ")
      p <- 1
      for (k in split){
        if (grepl("\\*", k) == TRUE){
          split[p] <- paste("\\", k, sep = "")
          print(split[k])
        }
        p <- p + 1
      }
      unique_mut <- stri_unique(split)
      for (j in unique_mut[-length(unique_mut)]){
        count <- str_count(i, j)
        mut_list_unique[pos] <- paste(mut_list_unique[pos], "<br>", toString(count), j)
        mut_list_unique_len[pos] <- mut_list_unique_len[pos] + 1
      }
      pos <- pos + 1
    }
    
    data_temp <- data.frame(x = seq(1, 566), y = mut_freq, list = mut_list_unique, list_len = mut_list_unique_len, group = df$group[1])
    groups <- c(groups, df$group[1])
    data_list <- c(data_list, list(data_temp))
  }
  
  all_false <- rep(FALSE, length(groups)*2)
  button_list <- list()
  for (i in 1:length(groups)) {
    visibility_state <- all_false
    visibility_state[(2 * i - 1):(i + i)] <- TRUE
    button_list[[i]] <- list(
      label = groups[i],
      method = "update",
      args = list(list(visible = visibility_state), list(title = groups[i]))
    )
  }
  
  p <- ggplot() +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank()
    ) +
    xlab("Amino Acid Position") +
    ylab("Number of samples") +
    xlim(c(0,566))
  
  for (i in 1:length(groups)) {
    d <- data_list[[i]][which(data_list[[i]]$y > 0),]
    row.names(d) <- seq(nrow(d))
    p <- p + 
      geom_segment(data = d, aes(x=x, xend=x, y=0, yend=y), color="darkgrey") +
      geom_point(
        data = d, 
        aes(
          x=x, y=y, 
          text=paste0(
          "<b>AA Position: </b>", x, 
          "<br><b># of alternative AA: </b>", y, 
          "<br><b>Substitutions: </b>", list_len, list
          )
        ),
        color="lightgrey",
        size=3
      )
  }
  
  ggplotly(p, tooltip = c("text")) %>% layout(
    updatemenus=list(
      list(
        active=0, 
        x=0, 
        y=1.1,
        xanchor="left",
        yanchor="bottom",
        buttons=button_list
      )
    )
  )
}
```

### Table

```{r grey table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(grey_table) == 0){
  cat("No sequences in Group GREY")
} else {
  for (i in grey_table_copy["Mutations"]){
    new <- gsub(",([[:alpha:]])", ", \\1", i)
    grey_table_copy["Mutations"][grey_table_copy["Mutations"] == i] <- new
  }
  
  subtype <- vector(mode = "character", length = length(grey_table_copy$ID))
  seg <- vector(mode = "character", length = length(grey_table_copy$ID))
    
  grey_table_copy <- cbind(subtype, seg, grey_table_copy)
    
  names(grey_table_copy)[1] <- "Subtype"
  names(grey_table_copy)[2] <- "Segment"
  
  for (j in 1:length(grey_table_copy$ID)){
    grey_table_copy$Subtype[j] <- str_split_i(grey_table_copy$ID[j], "\\|", -1)
    grey_table_copy$Segment[j] <- str_split_i(grey_table_copy$ID[j], "\\|", 2)
  }
  
  datatable(
    grey_table_copy, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 5,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      columnDefs = list(list(width = 500, targets = 4)),
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )
}
```

### Heatmap

```{r grey heatmap, out.width="1200px", echo=FALSE, message=FALSE, warning=FALSE}
if (!(nrow(grey_table) == 0)){
  names <- vector(mode = "character", length = length(grey_table$ID))
  df_names <- cbind(grey_table, names)
  names(df_names)[5] <- "Isolate_Name"
  for (j in 1:length(df_names$ID)){
    df_names$Isolate_Name[j] <- trimws(str_split_i(df_names$ID[j], "\\|", 3))
  }
}

if (nrow(grey_table) == 0){
  cat("No sequences in Group GREY")
} else if (params$metadata == ""){
  cat("The heatmap shows the frequency of mutations over the weeks.")
  cat("No metadata file with collection dates was given, so no heatmap was generated.")
} else if (any(df_names$Isolate_Name %in% metadata$Isolate_Name) == FALSE){
  cat("No metadata given for the sequences of Group GREY.")
} else {
  df_mutations_filtered <- merge(metadata, df_names, by="Isolate_Name") %>%
    select(Collection_Date, Mutations) %>%
    unnest(Mutations) %>%
    mutate(week = week(ymd(Collection_Date)))
    
  df_week <- df_mutations_filtered %>% 
    group_by(week) %>%
    summarize(count = n())
  
  df_mutations_filtered <- merge(df_mutations_filtered, df_week, by="week") %>%
    select(Mutations, week, count)

  names(df_mutations_filtered)[1] <- "aa_profile"
  names(df_mutations_filtered)[3] <- "week_count_total"

  df_mutations_grouped <- df_mutations_filtered %>% 
    group_by(week, week_count_total, aa_profile) %>% 
    summarize(freq = n()) %>%
    mutate(perc = round(freq/week_count_total*100, digits = 2)) %>% 
    ungroup() 
  
  pos <- c()
  for (i in df_mutations_grouped["aa_profile"]){
    # get position of mutation
    ind <- as.integer(str_sub(i, 2, -2))
    pos <- c(pos, ind)
  }
  df_mutations_grouped$position <- pos
  
  df_mutations_sorted <- df_mutations_grouped %>%
    arrange(-position)
  
  plot.data <- df_mutations_sorted %>% 
    select(week, aa_profile, freq, perc)
  
  min_val <- min(plot.data$week)
  max_val <- max(plot.data$week)
  full_seq <- min_val:max_val
  missing <- setdiff(full_seq, plot.data$week)
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], aa_profile = NA, freq = NA, perc = NA)
    plot.data <- rbind(plot.data, new_row)
  }
  
  aa_tick <- unique(plot.data$aa_profile)
  week_tick <- unique(plot.data$week)
  
  heatmap <- plot_ly(
    plot.data, x=~week, y=~aa_profile, z=~freq, 
    reversescale=TRUE, 
    xgap=1.5, ygap=1.5, 
    type="heatmap", 
    colorbar=list(
      title=('<b>Freq</b>')), 
      hoverinfo='text', 
      text=~paste0("<b>Week: </b>", .data$week, "<br><b>Substitution: </b>", .data$aa_profile, "<br><b># of Substitutions: </b>", .data$freq)
    ) %>% 
    layout(
      plot_bgcolor="#F8F8F8", 
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(nticks=length(aa_tick), title="<b>Substitutions</b>", showgrid=FALSE, categoryorder="trace"),
      legend=list(orientation='v', xanchor="right", yanchor="top")
    ) 
  
  bar.data <- metadata %>%
    select(Collection_Date) %>%
    mutate(week = week(ymd(Collection_Date))) %>%
    group_by(week) %>%
    summarize(count = n())
  
  for (i in 1:length(missing)){
    new_row <- list(week = missing[[i]], count = 0)
    bar.data <- rbind(bar.data, new_row)
  }
  
  bar.data <- bar.data %>% 
    arrange(week)
  
  barplot <- plot_ly(
    bar.data, x=~week, y=~count, 
    type = "bar",
    marker = list(color = "grey"),
    hoverinfo='text',
    hovertext=~paste0("<b>Week: </b>", .data$week, "<br><b># of Sequences: </b>", .data$count)
  ) %>%
    layout(
      xaxis=list(nticks=length(week_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(title="<b>Total Number <br> of Sequences</b>", showgrid=FALSE)
    )
  
  subplot(
    heatmap, barplot, 
    nrows = 2, heights = c(0.8, 0.2), 
    shareX = TRUE, titleY = TRUE
  )
}
```


<a id="item-five"></a>

## Group <span style="color: black;">BLACK</span>

```{r black data, echo=FALSE, message=FALSE, warning=FALSE}
black_table <- filter(samples, alert_level == 'black') 

cat(paste0('> ', 'Sequences in Group BLACK: ', nrow(black_table)))
```




<a id="item-six"></a>

## Reference sequences

The results from this report were calculated with the following reference sequence:

```{r ref, echo=FALSE, message=FALSE, warning=FALSE}
if (params$subtype == "h1n1" && params$ref == ""){
  cat("The Influenza A H1N1pdm reference sequence A/Wisconsin/588/2019 MW626065 was used.")
}else if (params$subtype == "h1n1" && params$ref == "old"){
  cat("The Influenza A H1N1pdm reference sequence A/California/7/2009 CY121680 was used.")
}else if (params$subtype == "h3n2" && params$ref == ""){
  cat("The Influenza A H3N2 reference sequenceA/Darwin/6/2021 EPI1857216 was used.")
}else if (params$subtype == "h3n2" && params$ref == "old"){
  cat("The Influenza A H3N2 reference sequence A/Wisconsin/67/2005 CY163680 was used.")
}else if (params$subtype == "vic" && params$ref == ""){
  cat("The Influenza B Victoria reference sequence B/Brisbane/60/2008 KX058884 was used.")
}
```

The annotations are generated with [Nextclade](https://clades.nextstrain.org). The following references are supported for the different Influenza subtypes.

#### Influenza A H1N1pdm09

For HA segment:

  - [A/Wisconsin/588/2019 MW626065](https://github.com/nextstrain/nextclade_data/blob/master/data/nextstrain/flu/h1n1pdm/ha/MW626062/reference.fasta)
  
  - [A/California/7/2009 CY121680](https://github.com/nextstrain/nextclade_data/blob/master/data/nextstrain/flu/h1n1pdm/ha/CY121680/reference.fasta)

#### Influenza A H3N2

For HA segment:

  - [A/Darwin/6/2021 EPI1857216](https://github.com/nextstrain/nextclade_data/blob/master/data/nextstrain/flu/h3n2/ha/EPI1857216/reference.fasta)

  - [A/Wisconsin/67/2005 CY163680](https://github.com/nextstrain/nextclade_data/blob/master/data/nextstrain/flu/h3n2/ha/CY163680/reference.fasta)

#### Influenza B Victoria 

For HA segment:

  - [B/Brisbane/60/2008 KX058884](https://github.com/nextstrain/nextclade_data/blob/master/data/nextstrain/flu/vic/ha/KX058884/reference.fasta)


<a id="item-seven"></a>

## MOC table

The following table shows a list of the Mutations of Concern that were used to generate the Report.

```{r moc, echo=FALSE, message=FALSE, warning=FALSE}
moc <- moc[, -4]

datatable(
  moc, 
  class = 'cell-border stripe', 
  options = list(
    dom = 'lftp', 
    lengthMenu = c(5, 10, 20, 50), 
    pageLength = 10,
    searchHighlight = TRUE,
    initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}")
  )
)
```


<a id="item-eight"></a>

## ROI table

The following table shows a list of Regions of Interest that were used to generate the Report.

```{r roi, echo=FALSE, message=FALSE, warning=FALSE}
names(roi)[1] <- "gene"

datatable(
  roi, 
  class = 'cell-border stripe', 
  options = list(
    dom = 'lftp', 
    lengthMenu = c(5, 10, 20, 50),
    pageLength = 10,
    searchHighlight = TRUE,
    initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}")
  )
)
```
